---
title: "Iris data Classification (SVM)"
author: "Piyush Pallav"
date: "26 February 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Here we will look at the famous iris Dataset and try to classify the its Species based on the given parameters. The dataset has been broken into two groups: *training set* and *test set* with split ratio being 80:20.

### Loading the dependent library and Dataset
```{r libraries, warning=FALSE}
library(caTools)
library(ggplot2)
library(GGally)
library(e1071)
```

We will use **caTools** for Dataset spliting into training & test set.

**ggplot2** and **GGally** will be used for Visualization.

**e1071** will be used for Support Vector Classification.

```{r dataset, warning=FALSE}
dataset = iris
```

### Exploring the dataset
```{r summary}
str(dataset)
summary(dataset)
```

We will use *sepal length*, *sepal width*, *petal length* and *petal width* to predict the *species* of Flower.

### Spliting data into training set and test set
```{r splitting}
split = sample.split(dataset$Species, SplitRatio = .8)
training_set = subset(dataset, split == TRUE)
test_set = subset(dataset, split == FALSE)

nrow(training_set)
nrow(test_set)
```

We have 120 data points on which we will train our model and then we will use 30 data points to test the model on.

### Exploratory Visualization

Lets have a closer look at the parameters and judge before hand if a good model can be created or not.

```{r visual, warning=FALSE}
ggpairs(training_set, ggplot2::aes(colour = Species, alpha = 0.4))
```

I personally like this graph because you can deduce so much information from one single chart. Lets have a look!

We can clearly see from the Histograms of *Petal.length* and *Petal.width* that we can clearly seperate out *Setosa* species with very high confidence.

However, *Versicolor* and *Virginica* Species are overlapped. If we look at the scatterplot of *Sepal.Length* vs *Petal.Length* and *Petal.Width* vs *Petal.Length*, we can distintly see a seperator that can be draw between the groups of Species. 

Looks like we can just use *Petal.Width* and *Petal.Length* as parameters and come with a good model. SVM seems to be a very good model for this type of data. Lets create two model, one contains all parameter and second contain just *Petal.Width* and *Petal.Length* as parameter and compare their individual performances. 

### Feature Scaling and Model Fiting

```{r feature scale}
training_set[,1:4] = scale(training_set[,1:4])
test_set[,1:4] = scale(test_set[,1:4])

classifier1 = svm(formula = Species~., data = training_set, type = 'C-classification', kernel = 'radial')
classifier2 = svm(formula = Species~ Petal.Width + Petal.Length, data = training_set, type = 'C-classification', kernel = 'radial')
```

Here classifier1 uses all the parameter to make model while *classifier2* just uses Petal's Legth and Width to generate model.

### Prediction and Evaluation

```{r pred}
test_pred1 = predict(classifier1, type = 'response', newdata = test_set[-5])
test_pred2 = predict(classifier2, type = 'response', newdata = test_set[-5])

# Making Confusion Matrix
cm1 = table(test_set[,5], test_pred1)
cm2 = table(test_set[,5], test_pred2)
cm1 # Confusion Matrix for all parameters
cm2 # Confusion Matrix for parameters being Petal Length and Petal Width
```

Wow, the accuracy for both model looks solid. Also notice that as we had deduced, only *Petal* Length and Width is important to make this model accurate and our second classifier proves it!